{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-12-07 ident_manu.tex
personal edits to results, discussion

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Tue Dec 6 13:56:23 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) evaluated parameters in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton), whereas average sensitivity among all parameters was $L1 = $ $0.01$. Within categories (excluding temperature), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $0.01$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $0$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All variables were sensitive to the same parameters as \ac{do} (38 of 51 evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from $0.02$ for irradiance (\cref{tab:irrsens}) to $0.71$ for \ac{dom} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = $0.01$).  For each variable, $L1$ ranged from $2.24\times 10^{-6}$ (\textit{QminP}) to $8.49$ (\textit{mA}) for ammonium (\cref{tab:nh4sens}), $1.38\times 10^{-6}$ (\textit{QminP}) to $13.94$ (\textit{mA}) for \ac{chla} (\cref{tab:chlsens}), $1.92\times 10^{-7}$ (\textit{QminP}) to $0.13$ (\textit{ZKa}) for irradiance (\cref{tab:irrsens}), $6.67\times 10^{-7}$ (\textit{QminP}) to $8.49$ (\textit{umax}) for nitrate (\cref{tab:no3sens}), $6.41\times 10^{-5}$ (\textit{KNH4}) to $7.22$ (\textit{mA}) for \ac{pom} (\cref{tab:om1sens}),  $7.41\times 10^{-5}$ (\textit{KNH4}) to $14.25$ (\textit{mA}) for \ac{dom} (\cref{tab:om2sens}), and $8.21\times 10^{-7}$ (\textit{QminP}) to $1.47$ (\textit{ZKa}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to phytoplankton parameters (average $L1$ = $0.8$ across all parameters in the category), \ac{chla} to phytoplankton ($L1$ = $1.14$), irradiance to zooplankton ($L1$ = $0.03$), nitrate to zooplankton ($L1$ = $1.06$), \ac{pom} to temperature ($L1$ = $0.86$), \ac{dom} to temperature ($L1$ = $1.48$), and phosphate to zooplankton ($L1$ = $0.31$).  Finally, average sensitivity between parameter categories independent of the state variables ranged from $0.01$ for optics (average $L1$ across all variables) to $0.62$ for phytoplankton.

@@

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Results for \ac{do} are provided first to demonstrate general concepts for the identifiability analyses, followed by an extension to the remaining state variables.  Parameter identifiability (as $\gamma$) for \ac{do} increased at different rates with increasing size of parameter subsets depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets in the phytoplankton ($60$\% of subsets less than $\gamma = 15$, $43$\% less than $\gamma = 10$) and zooplankton categories ($53.1$\% less than $\gamma = 15$, $40$\% less than $\gamma = 10$), whereas all combinations were identifiable for optics ($100$\% less than $\gamma = 15$, $100$\% less than $\gamma = 10$) and a majority identifiable for organic matter ($91.9$\% less than $\gamma = 15$, $76.5$\% less than $\gamma = 10$). Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was $100$\% less than $\gamma = 15$ and $100$\% less than $\gamma = 10$ for the top parameters in each category, $90.6$\% and $80.7$\% for the top two, $80.7$\% and $70.9$\% for the top three, and $55.8$\% and $45.7$\% for the fop four.  Results for the remaining state variables had similar patterns in identifiability with increasing size of parameter subsets and selection categories, although differences in identifiabiltiy between state variables was observed (\cref{fig:identploall}).  Most notably, nitrate was consistently the least identifiable variable for all selection heuristics (highest $\gamma$), whereas \ac{do} was most identifiable.

An alternative view of the results in \cref{fig:identplo} can be used to demonstrate the effects of parameter selection criteria and number of parameters in the selection pool on identifiability.  \Cref{fig:percthresh} shows the percentage of identifiable parameter sets for \ac{do} using the same selection criteria in \cref{fig:identplo}, i.e., selection of parameters only within parameter categories and selection of the top sensitive parameters regardless of category.  \Cref{fig:percthresh} is conceptually similar to \cref{fig:identplo}, with the added effect of a chosen $\gamma$ threshold on identifability.  Previous studies have provided general rules for $\gamma$ thresholds \citep{Brun01,Omlin01}, such that exact values beyond which model calibration fails could vary for the given model and optimization method.  The selected $\gamma$ thresholds are somewhat arbitrary, although the results show differences that have implications for parameter selection for model calibration.  In general, identifiability decreased with the addition of parameters, although the rate of decrease depended on the selected threshold for $\gamma$. More conservative values for $\gamma$ (e.g., $\gamma = 5$) were more sensitive to the number of parameters in a subset, that is, identifiability decreased more quickly with the addition of parameters at lower $\gamma$ thresholds as compared to higher $\gamma$.  Notable differences in identifiability were also observed by selection criteria (within categories or top parameters only), which further supports results in \cref{fig:identplo,fig:identploall}.

@@

The above results demonstrated that the state variables differed in the magnitudes of sensitivity for each parameter and the number of identifiable subsets, where the latter varied by identifiability thresholds and parameter selection criteria.  These results further underscore the need for explicit selection heuristics for parameters to calibrate for improving model performance. Results for each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) applied to each state variable differed in the number of selected parameters and distribution of parameters within each category (\cref{tab:heurist1,tab:heurist2,tab:heurist3}).  In general, a corresponendence was observed between the number of parameters that were selected given the threhold of $\gamma = 15$ and relative identifiability between the state variables.  As noted above, nitrate was the least identifiable variable (\cref{fig:identploall}), whereas other variables (e.g., \ac{do}, irradiance) were more identifiable.  The constraints on relative identifiability between variables were demonstrated with the selection heuristics.  For example, heuristics for nitrate typically selected only one or two parameters that met the criteria as compared to more identifiable variables that included several parameters that were identifiable. Overall, the first selection heuristic demonstrated that the number of parameters chosen by parameter category differed independently of the state variables (\cref{tab:heurist1}). The number of selected parameters averaged across state variables in decreasing order was 4.25 parameters from the phytoplankton category, 3.5 from organic matter, 2.75 from optics, and 2.38 from zooplankton.  These rankings were independent of parameter sensitivity and identifiability in each category.  The second and third selection heuristics (\cref{tab:heurist2,tab:heurist3}) were similar, although more parameters were generally selected for the third heuristic.

\Cref{fig:heurist_stts} demonstrates parameter selection for all state variables following the second heuristic where parameters were chosen by decreasing sensitivity independent of parameter categories (exact values are shown in \cref{tab:heurist2}).  The y-axis on the plot shows the relative identifiability values with the addition of parameters from one to many on the x-axis (from left to right).  The second to last parameter for each variable is the last parameter selected within the potential threshold of $\gamma = 15$.  Interestingly, the last parameter shown for most of the state variables caused a relatively large increase in $\gamma$ that was disproportionate to the combined identifiability of the preceding parameter set. For most variables, the phytoplankton edibility vector for zooplankton (\textit{ediblevector(Z1)}) caused a dramatic increase in $\gamma$ with inclusion in the parameter set.  In addition to demonstrating the approach for selection with the second heuristic, \cref{fig:heurist_stts} provides an alternative method of identifying parameters that are disporportionately redundant within a given parameter set.  This infomation is useful if additional parameter selection rules are developed independently of those proposed herein.

@@

Both the relative $L1$ estimates and number of parameters that affected state variables were sensitive to variation in the initial conditions (observation effects) and structural changes in the model.  Visual comparison of sensitivity values ($L1$) for the default model setup showed that all state variables were sensitive to \textit{vmaxSi} after changing the initial conditions to April or September conditions (\cref{fig:sensalltile}).  State variables were also sensitive to \textit{Ksi} for only the September conditions. Changing the model structure from the default (simple) to the complex setup showed state variables were also sensitive to \textit{KQn} and \textit{Tref(nospA+nospZ)$_{z1}$}.  Changes in the sensitivity of individual parameters were also observed, most notably as a disproportionate increase in $L1$ for ammonium, \ac{chla}, \ac{pom}, and \ac{dom} to \textit{mA} and \textit{Tref(nospA+nospZ)$_{p1}$} using the complex model setup.  Summaries by parameter categories of the effects of initial conditions and structural changes are shown in \cref{tab:inistrchg,fig:inistrsumm}. Changes in average sensitivity using April and September initial conditions showed most variables were less sensitive to parameters in the optics, temperature, and zooplankton categories, whereas increased sensitivity was generally observed for parameters in the organic matter and phytoplankton categories.  Some exceptions were observed, such as a decrease in nitrogen sensitivity to phytoplankton parameters.  Effects of structural complexity were most apparent as an increase in sensitivity of state variables to phytoplankton and temperature parameters, as noted for the individual variables in \cref{fig:sensalltile}.

@@

Common goals in the application of biogeochemical models of ecosystem processes are to \begin{inparaenum}[1\upshape)]
\item accurately describe the system my matching predictions with observed data \citep{Reckhow90}, and 
\item provide a means of forecasting ecosystem condition with hypothesized management or environmental scenarios \citep{Clark01}. 
\end{inparaenum}
Although these objectives are the focus of most model applications, the structural components of process-based models should secondarily provide inference into which ecosystem processes and functions are driving observed or hypothesized changes.  This latter objective represents a more general scientific outcome of biogeochemical models that extends beyond the applied benefits of describing and predicting change in a particular system.  Modelers often hope to identify universal principles that govern dynamics across systems and the constraint of model parameters to observations provides a means of supporting or refuting hypotheses \citep{Kirchner06}.  Extension of these principles to test the effects of structural changes and observation uncertainty on model predictions provides further evidence to validate model components. This study provided a simple approach to use the effects of parameter perturbations on model state variables to characterize identifiable parameter subsets that vary by parameter selection criteria.  By doing so, we demonstrated that small parameter subsets relative to all sensitive parameters were within the identifiabilty thresholds described in the literature.  The identifiable parameter subsets varied considerably between state variables and the method for parameter selection.  We further demonstrated that changes in the model structure and variation in the initial conditions had an effect on sensitivity which has direct implications for identifiability. In general, these results provide justification for the use of explicit parameter selection heuristics that practitioners should adopt for model calibration.

Although the results were specific to individual variables, some generalities can be inferred from the sensitivity analyses. State variables were most sensitive to parameters in the phytoplankton and zooplankton categories, particularly the maximum growth rates (\textit{umax} for phytoplankton, \textit{Zumax} for zooplankton), mortality coefficient for phytoplankton (\textit{mA}), and the zooplankton half saturation coefficent for grazing (\textit{ZKa}). An increase in growth rate of primary producers has the potential to increase oxygen concentration through photosynthetic processes, although increased production of organic matter is balanced with respiration and bacterial decomposition that reduce \ac{do} concentration in the water column.  Similarly, increases in zooplankton abundance with increased growth rates causes a reduction in phytoplankton biomass through grazing, which is expected to further deplete pools of organic matter.  Most variables were also sensitive to variation in the half-saturation grazing coefficient which moderates the concentration of nutrients that support half the maximum grazing rate. Although the tradeoff between abundance, grazing, and decomposition is complex, the sensitivity of model state variables to parameters that directly control the abundance of primary producers is in agreement with empirical observations of factors that influence hypoxia dynamics on the \ac{lcs} \citep{Fahnenstiel95,Roelke00,Eldridge10}. The sensitivity of the model output to variation in other parameters that relate to physical and chemical properties of the system was of secondary importance to biological relationships.  State variables were sensitive to changes in light and temperature parameters, although to a lesser extent than phytoplankton and zooplankton parameters.  As such, the differing sensitivities of state variables to parameters in each of the categories was not unexpected given general ecological relationships that are well understood and described by the model.

The overwhelming conclusion from the identifiablity analyses is that only limited subsets of parameters are identifiable within the constraints of local sensitivity analyses. Although we have not attempted actual model calibration (see recommendations below), these results support previous studies that have suggested similarly small subsets of parameters can be identified using traditional calibration schemes \citep[e.g.,][]{Wheater86,Ye97,Omlin01}.  In addition to \ac{cgem}, these conclusions have relevance for many biogeochemical models that include numerous parameters and structural equations to characterize processes in the model domain.  A general conclusion is that perhaps a trend towards less complex models could be beneficial given that only a small subset of parameters is identifiable and that ecosystem processes may in fact be sufficiently characterized with few parameters \citep{Ye97}.  Conversely, others have argued that model complexity is not in itself a disadvantage when parsimony is not the sole arbitrator of model structure \citep{Reichert97}. Over-parameterization can be useful if processes have importance that were not evaluated during model identification.  Single objective functions that maximize model precision with identifiable parameters may also provide an incomplete characterization of model worth, which has prompted the development of probability-based models of hypoxia that explicitly include uncertainty in model components \citep[e.g.][]{Obenour15}.  Our results demonstrated that approximately 75\% of the evaluated parameters had an effect on the eight state variables, whereas \ac{cgem} includes a total of 36 variables and multiple plankton groups, not all of which have immediate concern for understanding hypoxia. The redundancies identified with the sensitivity analyses are only problematic if the primary interest is, for example, \ac{do} dynamics.  Moreover, the proposed parameter selection heuristics provide flexibility for selecting different parameters based on parameter categories with the assumption that the chosen parameters depends on the research or management question.

>>>>

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Wed Dec 7 11:57:48 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) parameters that were evaluated in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton), whereas average sensitivity among all parameters was $L1 = $ $9.2\times 10^{-3}$. Within categories (excluding temperature with one sensitive parameter), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $6.15\times 10^{-3}$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $2.17\times 10^{-3}$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

Local sensitivity analyses for the additional state variables (ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus) had similar results as \ac{do} with some exceptions (\cref{fig:sensalltile,tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens}).  All variables were sensitive to the same parameters as \ac{do} (38 of 51 evaluated), although average sensitivity differed between variables.  Average $L1$ ranged from $0.02$ for irradiance (\cref{tab:irrsens}) to $0.71$ for \ac{dom} (\cref{tab:om2sens}).  All average sensitivity values for the state variables were higher than the average for \ac{do} ($L1$ = $9.2\times 10^{-3}$).  For each variable, $L1$ ranged from $2.24\times 10^{-6}$ (\textit{QminP}) to $8.49$ (\textit{mA}) for ammonium (\cref{tab:nh4sens}), $1.38\times 10^{-6}$ (\textit{QminP}) to $13.94$ (\textit{mA}) for \ac{chla} (\cref{tab:chlsens}), $1.92\times 10^{-7}$ (\textit{QminP}) to $0.13$ (\textit{ZKa}) for irradiance (\cref{tab:irrsens}), $6.67\times 10^{-7}$ (\textit{QminP}) to $8.49$ (\textit{umax}) for nitrate (\cref{tab:no3sens}), $6.41\times 10^{-5}$ (\textit{KNH4}) to $7.22$ (\textit{mA}) for \ac{pom} (\cref{tab:om1sens}),  $7.41\times 10^{-5}$ (\textit{KNH4}) to $14.25$ (\textit{mA}) for \ac{dom} (\cref{tab:om2sens}), and $8.21\times 10^{-7}$ (\textit{QminP}) to $1.47$ (\textit{ZKa}) for phosphate (\cref{tab:po4sens}).  For the parameter categories, ammonium was most sensitive to phytoplankton parameters (average $L1$ = $0.8$ across all parameters in the category), \ac{chla} to phytoplankton ($L1$ = $1.14$), irradiance to zooplankton ($L1$ = $0.03$), nitrate to zooplankton ($L1$ = $1.06$), \ac{pom} to temperature ($L1$ = $0.86$), \ac{dom} to temperature ($L1$ = $1.48$), and phosphate to zooplankton ($L1$ = $0.31$).  Finally, average sensitivity between parameter categories independent of the state variables ranged from $8.38\times 10^{-3}$ for optics (average $L1$ across all variables) to $0.62$ for phytoplankton.

@@

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Results for \ac{do} are provided first to demonstrate general concepts for the identifiability analyses, followed by an extension to the remaining state variables.  Parameter identifiability for \ac{do} decreased (increasing $\gamma$) at different rates with increasing size of parameter subsets depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets in the phytoplankton ($60$\% of subsets less than $\gamma = 15$, $43$\% less than $\gamma = 10$) and zooplankton categories ($53.1$\% less than $\gamma = 15$, $40$\% less than $\gamma = 10$), whereas all combinations were identifiable for optics ($100$\% less than $\gamma = 15, 10$) and a majority identifiable for organic matter ($91.9$\% less than $\gamma = 15$, $76.5$\% less than $\gamma = 10$). Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was $100$\% less than $\gamma = 15, 10$ for the top parameter in each category, $90.6$\% and $80.7$\% for the top two, $80.7$\% and $70.9$\% for the top three, and $55.8$\% and $45.7$\% for the top four.  Results for the remaining state variables had similar patterns in identifiability with increasing size of parameter subsets and selection categories, although differences in identifiabiltiy between state variables was observed (\cref{fig:identploall}).  Most notably, nitrate was consistently the least identifiable variable (highest overall $\gamma$), whereas \ac{do} was most identifiable.

An alternative view of the results in \cref{fig:identplo} can be used to demonstrate the effects of parameter selection criteria and number of parameters in the selection pool on identifiability.  \Cref{fig:percthresh} shows the percentage of identifiable parameter sets for \ac{do} using the same selection criteria in \cref{fig:identplo}, i.e., selection of parameters only within parameter categories and selection of the top sensitive parameters regardless of category.  \Cref{fig:percthresh} is similar to \cref{fig:identplo}, with the added effect of a chosen $\gamma$ threshold on identifability.  Previous studies have provided only general rules for $\gamma$ thresholds \citep{Brun01,Omlin01}, such that exact values for which parameter sets are inestimable likely vary between models and optimization methods.  As such, multiple values are shown in \cref{fig:percthresh}  given that actual thresholds could vary in practice. In general, identifiability decreased with the addition of parameters, although the rate of decrease depended on the selected threshold for $\gamma$. More conservative values for $\gamma$ (e.g., $\gamma = 5$) were more sensitive to the number of parameters in a subset, that is, identifiability decreased more quickly with the addition of parameters at lower $\gamma$ thresholds as compared to higher $\gamma$.  Notable differences in identifiability were also observed by parameter selection criteria (within categories or top parameters only), which further supports results in \cref{fig:identplo,fig:identploall}.

@@

The above results demonstrated that state variables differed in the magnitudes of sensitivity for each parameter and the number of identifiable subsets, where the latter varied by $\gamma$ thresholds and parameter selection criteria. Results for each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) applied to each state variable differed in the number of selected parameters and distribution of parameters within each category (\cref{tab:heurist1,tab:heurist2,tab:heurist3}).  In general, a corresponendence was observed between the number of parameters that were selected given the threhold of $\gamma = 15$ and relative identifiability between the state variables.  As noted above, nitrate was the least identifiable variable (\cref{fig:identploall}), whereas other variables (e.g., \ac{do}, irradiance) were more identifiable.  The constraints on identifiability between variables were demonstrated with the selection heuristics.  For example, heuristics for nitrate typically selected only one or two parameters that met the criteria as compared to more identifiable variables that included several parameters. Overall, the first selection heuristic demonstrated that the number of parameters chosen by parameter category differed independently of the state variables (\cref{tab:heurist1}). The number of selected parameters averaged across state variables in decreasing order was 4.25 parameters from the phytoplankton category, 3.5 from organic matter, 2.75 from optics, and 2.38 from zooplankton. The second and third selection heuristics (\cref{tab:heurist2,tab:heurist3}) were similar, although more parameters were generally selected for the third heuristic given equal importance between categories.

\Cref{fig:heurist_stts} demonstrates parameter selection for all state variables following the second heuristic where parameters were chosen by decreasing sensitivity independent of parameter categories (exact values are shown in \cref{tab:heurist2}).  The y-axis shows the relative identifiability with the addition of parameters from one to many on the x-axis (from left to right).  The second to last parameter for each variable is the last parameter selected within the potential threshold of $\gamma = 15$.  Interestingly, the last parameter shown for most of the state variables caused a relatively large increase in $\gamma$ that was disproportionate to the combined identifiability of the preceding parameter set. For most variables, the phytoplankton edibility vector for zooplankton (\textit{ediblevector(Z1)}) caused a dramatic increase in $\gamma$ with inclusion in the parameter set.  In addition to demonstrating the approach for selection with the second heuristic, \cref{fig:heurist_stts} provides an alternative method of identifying parameters that are disproportionately redundant within a given parameter set (i.e., large contribution to $\gamma$).  This infomation is useful if additional parameter selection rules are developed independently of those proposed herein.

@@

Both the relative $L1$ estimates and number of parameters that affected state variables were sensitive to variation in the initial conditions (observation effects) and structural changes in the model.  Visual comparison of sensitivity values ($L1$) for the default model setup showed that all state variables were sensitive to \textit{vmaxSi} after changing the initial conditions to April or September conditions (\cref{fig:sensalltile}).  State variables were also sensitive to \textit{Ksi} for only the September conditions. Changing the model structure from the default (simple) to the complex setup showed state variables were also sensitive to \textit{KQn} and \textit{Tref(nospA+nospZ)$_{z1}$}.  Changes in the sensitivity of individual parameters were also observed, most notably as a disproportionate increase in $L1$ for ammonium, \ac{chla}, \ac{pom}, and \ac{dom} to \textit{mA} and \textit{Tref(nospA+nospZ)$_{p1}$} using the complex model setup.  Summaries by parameter categories of the effects of initial conditions and structural changes are shown in \cref{tab:inistrchg,fig:inistrsumm}. Changes in average sensitivity using April and September initial conditions showed most variables were less sensitive to parameters in the optics, temperature, and zooplankton categories, whereas increased sensitivity was generally observed for parameters in the organic matter and phytoplankton categories.  Some exceptions were observed, such as a decrease in nitrogen sensitivity to phytoplankton parameters (\cref{tab:inistrchg}).  Effects of structural complexity were most apparent as an increase in sensitivity of state variables to phytoplankton and temperature parameters, as noted for the individual variables in \cref{fig:sensalltile}.

@@

Common goals in the application of biogeochemical models of ecosystem processes are to \begin{inparaenum}[1\upshape)]
\item accurately describe the system by matching predictions with observed data \citep{Reckhow90}, and 
\item provide a means of forecasting ecosystem condition with hypothesized management or environmental scenarios \citep{Clark01}. 
\end{inparaenum}
Although these objectives are the focus of most applications, the structural components of process-based models should secondarily provide inference into which ecosystem processes and functions are driving observed changes.  This latter objective represents a more generic scientific purpose of biogeochemical models that extends beyond the applied benefits of describing and predicting change in a particular system.  Modelers often hope to identify universal principles that govern dynamics across systems and the constraint of model parameters to observations provides a means of supporting or refuting hypotheses \citep{Kirchner06}.  Extension of these principles to test the effects of structural changes and observation uncertainty on model predictions provides further information to support validation of model components. This study provided a simple approach to use the effects of parameter perturbations on model state variables to characterize identifiable parameter subsets that vary by parameter selection criteria.  By doing so, we demonstrated that small parameter subsets relative to all sensitive parameters were within the identifiabilty thresholds described in the literature.  The identifiable parameter subsets varied considerably between state variables and the method for parameter selection.  We further demonstrated that changes in the model structure and variation in the initial conditions had an effect on sensitivity which has direct implications for identifiability. In general, these results provide justification for the use of explicit parameter selection heuristics that practitioners should adopt to facilitate model calibration.

Although the results were specific to individual variables, some generalities can be inferred from the sensitivity analyses. State variables were most sensitive to parameters in the phytoplankton and zooplankton categories, particularly the maximum growth rates (\textit{umax} for phytoplankton, \textit{Zumax} for zooplankton), mortality coefficient for phytoplankton (\textit{mA}), and the zooplankton half saturation coefficent for grazing (\textit{ZKa}). An increase in the growth rate of primary producers has the potential to increase oxygen concentration through photosynthetic processes, although increased production of organic matter is balanced with respiration and bacterial decomposition that reduce \ac{do} in the water column.  Similarly, increases in zooplankton abundance with increased growth rates causes a reduction in phytoplankton biomass through grazing, which is expected to further deplete pools of organic matter.  Most variables were also sensitive to variation in the half-saturation grazing coefficient which moderates the concentration of nutrients that support half the maximum grazing rate. Although the tradeoff between abundance, grazing, and decomposition is complex, the sensitivity of model state variables to parameters that directly control the abundance of primary producers is in agreement with empirical observations of factors that influence hypoxia dynamics on the \ac{lcs} \citep{Fahnenstiel95,Roelke00,Eldridge10}. The sensitivity of the model output to variation in other parameters that relate to physical and chemical properties of the system was of secondary importance to biological relationships.  That is, state variables were sensitive to changes in light and temperature parameters, although to a lesser extent than phytoplankton and zooplankton parameters.  As such, the differing sensitivities of state variables to parameters in each of the categories was not unexpected given general ecological relationships that are well understood and described by the model.

The overwhelming conclusion from the identifiablity analyses is that only limited subsets of parameters are identifiable within the constraints of local sensitivity analyses. Although we have not attempted actual model calibration (see recommendations below), these results support previous studies that have suggested similarly small subsets of parameters can be identified using traditional calibration schemes \citep[e.g.,][]{Wheater86,Ye97,Omlin01}.  In addition to \ac{cgem}, these conclusions have relevance for many biogeochemical models that include numerous parameters and structural equations to characterize processes in the model domain.  A general conclusion is that perhaps a trend towards less complex models could be beneficial given that only a small subset of parameters is identifiable and that ecosystem processes may in fact be sufficiently characterized with few parameters \citep{Ye97}.  Conversely, others have argued that model complexity is not in itself a disadvantage when parsimony is not the sole arbitrator of model structure \citep{Reichert97}. Over-parameterization can be useful if processes have importance that were not evaluated during model identification.  Single objective functions that maximize model precision with identifiable parameters may also provide an incomplete characterization of model worth, which has prompted the development of probability-based models of hypoxia that explicitly include uncertainty in model components \citep[e.g.][]{Obenour15}.  Our results demonstrated that approximately 75\% of the evaluated parameters had an effect on the eight state variables, whereas \ac{cgem} includes a total of 36 variables and multiple plankton groups, not all of which have immediate concern for understanding hypoxia. The redundancies identified with the sensitivity analyses are only problematic if the primary interest is, for example, \ac{do} dynamics.  Moreover, the proposed selection heuristics provide flexibility for choosing different parameters with the assumption that those chosen depend on the research or management question.

Results from the identifiability analyses provided additional insight into the interactions of parameters in large biogeochemical models.  First, identifiability of parameter subsets was not related to the sensitivity of individual variables. As noted above, an identifiable parameter is one that has a unique effect on model predictions that cannot be compensated for or undone by changing other parameters. The magnitude of the effect of a parameter has no bearing on identifiability, which further complicates the selection of parameters for calibration.  Although identifiability is the primary limiting factor in choosing a set, the relative sensitivities are more important for the decision to include or exclude indvidual parameters.  Our analysis addressed this challenge by presenting multiple selection criteria for identifiable parameter sets that prioritized the most sensitive parameters during the selection process.  Similarly, identifability was not always related to the number of parameters in a set. Although the overwhelming trend was decreasing identifiability with more parameters, the unique effects of including an individual parameter with an existing set often reduced the $\gamma$ estimate. For example, \cref{fig:exclex} shows that including \textit{KGcdom}, \textit{KO2}, or \textit{nitmax} in parameter sets more often reduced $\gamma$ relative to sets that excluded the parameters.  Conversely, \cref{fig:heurist_stts} showed that the inclusion of a specific parameters often caused a disproportionate increase in $\gamma$ (e.g., \textit{ediblevector(Z1)}). These examples demonstrate the complex interactions of parameter changes on variable response, highlighting the need to consider the combined and individual effects of parameters on identifiability. The selection criteria proposed in our above analyses can facilitate parameter selection and also provide diagnostic tools to identify parameters with disproportionate effects on $\gamma$.
